---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
require(knitr)
opts_chunk$set(echo = TRUE, results = "markup", message = FALSE)
Sys.setlocale("LC_TIME", "English")
```

## Loading and preprocessing the data
```{r load_preprocess, warning = FALSE}
require(dplyr)
require(zoo)
require(lattice)
unzip("activity.zip")
activity <- read.csv("activity.csv", stringsAsFactors = FALSE)
activity$date = as.Date(activity$date)
```

## What is mean total number of steps taken per day?

### Histogram
```{r histogram}
stepsByDay <-
    activity %>%
    group_by(date) %>%
    summarise(tot = sum(steps))
hist(stepsByDay$tot,breaks=10)
```

### Mean value
```{r mean}
mean(stepsByDay$tot, na.rm = TRUE)
```

### Median value
```{r median}
median(stepsByDay$tot, na.rm = TRUE)
```

## What is the average daily activity pattern?

### Plot of typical day
```{r plot_daily}
stepsByInterval <-
    activity %>%
    group_by(interval) %>%
    summarise(avg = mean(steps, na.rm = TRUE))
plot(stepsByInterval, type = "l")
```

### Most intense interval
The following is the 5-minute interval containing the maximum number of steps:
```{r intense_interval}
stepsByInterval$interval[which.max(stepsByInterval$avg)]
```

## Imputing missing values

### Number of missing values
```{r nbr_na}
sum(is.na(activity$steps))
```

### Filling missing values
To fill NA values my strategy is repetition of the leftmost or rightmost non-NA value in the exterior, and linear interpolation in the interior. I do this using the function na.fill from the zoo package:
```{r na_fill}
activityFilled <- activity
activityFilled$steps <- na.fill(activity$steps, "extend")
stepsByDayFilled <-
    activityFilled %>%
    group_by(date) %>%
    summarise(tot = sum(steps))
hist(stepsByDayFilled$tot,breaks=10)
```

### Mean value
```{r mean2}
mean(stepsByDayFilled$tot)
```

### Median value
```{r median2}
median(stepsByDayFilled$tot)
```

The values are now lower, as my strategy gives more days with zero activity.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays}
day <- as.factor(ifelse(weekdays(activityFilled$date) %in% c("Saturday","Sunday"), "Weekend", "Weekday"))
activityExtended <- cbind(activityFilled, day)
stepsByIntervalExtended <-
    activityExtended %>%
    group_by(interval, day) %>%
    summarise(avg = mean(steps, na.rm = TRUE))
xyplot(avg ~ interval|day,stepsByIntervalExtended, type = "l", layout = c(1,2))
```